---
title: "WQD7004 2024/2025 Sem2 - Group G03"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
date: "2025-06-01"
---

<style>
body {
  background-color: #f7cac9;
}
</style>

::: {style="text-align: right;"}
Group G03

Richie Teoh (24088171)\
Elmer Lee (2408236)\
Micole ()\
Lee Re Xuan ()\
Angeline Tan Jie Lin ()

01 June 2025
:::

------------------------------------------------------------------------

# **1 Objectives**

Title: 

- Prediction of Rental Price via Regression and Classification of Property Furnishing Level in Klang Valley, Malaysia.

Project Goal:

- There are two (2) goals we wish to achieve with this project:
1) *Prediction of Rental Prices*: We aim to estimate the rental price of a property based on factors like location, size, amenities. An accurate estimate will help potential renters such as students and workers to get a fair market value.

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
```

```{r}
#setwd('C:/Users/Richie.teoh/Downloads/UM/WQD7004/assignment')
df <- read.csv('mudah-apartment-kl-selangor.csv')
df
```

```{r}
metadata <- tibble(
  Column_Name = c(
    "ads_id",
    "prop_name",
    "completion_year",
    "monthly_rent",
    "location",
    "property_type",
    "rooms",
    "parking",
    "bathroom",
    "size",
    "furnished",
    "facilities",
    "additional_facilities"
  ),
  Description = c(
    "Unique identifier for the listing",
    "Name of the building/property",
    "Completion/established year of the property",
    "Monthly rent in Ringgit Malaysia (RM)",
    "Property location within Kuala Lumpur region",
    "Type of property (apartment, condo, flat, etc.)",
    "Number of rooms in the unit",
    "Number of parking spaces for the unit",
    "Number of bathrooms in the unit",
    "Total area of the unit in square feet",
    "Furnishing status of the unit (fully, partial, non-furnished)",
    "Main facilities available",
    "Additional facilities (e.g., proximity to mall, school, railway, etc.)"
  )
)

# Print the metadata table
print(metadata)
```
From the metadata extracted from Kaggle above:
- ads_id should be dropped as it is a unique identifier for the Mudah rental listing.
- prop_name can influence the rental price, for example properties under well-known developers such as Arte tend to be higher-priced; however, it does not generalize well and unseen property names might cause issues in the model. Henceforth, we decide to drop this variable as well.
- completion_year includes some NA values, will investigate further, completion_year contributes to the price and the info are usually available online, hence we will impute it manually from data found online.
- monthly_rent is saved as chr type due to the "RM" prefixes, will perform appropriate data transformation.
- size is saved as chr type due to the "sq. ft." suffixes, will perform appropriate data transformation.

```{r}
# Data Pre-processing

# Keep only serviced residences and condo
df <- df %>%
  filter(.data[['property_type']] %in% c('Service Residence','Condominium'))
# Check and Remove duplicated values.
duplicates_only <- df %>%
  filter(duplicated(df) | duplicated(df, fromLast = TRUE)) %>%
  arrange(ads_id)

duplicates_ads_id <- df %>%
  filter(duplicated(df['ads_id']) | duplicated(df['ads_id'], fromLast = TRUE)) %>%
  arrange(ads_id)

duplicates_only
duplicates_ads_id
```
Interestingly, despite *ads_id* being the unique identifier for the rental listing, duplicated(df) does not yield the same result as duplicated(df['ads_id']). Hence, we will only keep the row that contains the most information, or the first one should both rows contain the same amount of information.

```{r}
# Removing Duplicates
cleaned_df <- df %>%
  mutate(na_count = rowSums(is.na(.))) %>%
  group_by(ads_id) %>%
  arrange(na_count) %>%
  slice(1) %>%
  ungroup() %>%
  select(-na_count)

sum(duplicated(cleaned_df['ads_id']))
df <- cleaned_df

# 
```
## Check for Missing Data and Imputation
```{r}
sum(is.na(df))
colSums(is.na(df))
df[is.na(df['completion_year']),]
unique(df[is.na(df['completion_year']),]$prop_name)
```


## Data Transformation
```{r}
# Data Transformation

# 1) Removing prefix 'RM' and suffix 'per month' from monthly rent, removing white spaces from 1 700, and convert to numeric type.
df <- df %>%
  mutate(
    monthly_rent = str_remove_all(monthly_rent, "RM|per month|,"),
    monthly_rent = str_remove_all(monthly_rent, " "),
    monthly_rent = as.numeric(monthly_rent)
  )

# 2) Removing suffix 'sq. ft.' from size, removing any white spaces, and convert to numeric type.
df <- df%>%
  mutate(
    size = str_remove_all(size, "sq\\.ft\\."),
    size = str_remove_all(size, " "),
    size = as.numeric(size)
  )

unique(df[is.na(df$completion_year), "prop_name"])
df[is.na(df$monthly_rent),]

df <- df %>%
  filter(!is.na(monthly_rent))
```
Interesting case, Affina Bay is located in Butterworth but included in this webscrapped Klang Valley Mudah listing.
Henceforth, we will take more time to clean the data properly based on the listing's location.

```{r}
unique(df['location'])
```
Going through the unique values, we see a few interesting ones such as Selangor - 389, Selangor - 360, Kuala Lumpur - Others, Selangor - 517, Selangor - 639

```{r}
suspicious_location <- c('Selangor - 389', 'Selangor - 360', 'Kuala Lumpur - Others', 'Selangor - 517', 'Selangor - 639')

df[df$location %in% suspicious_location,]

df %>%
  filter(grepl("Ayuman", prop_name))
```
```{r}
# Investigating number of rooms
unique(df$rooms)

df <- df %>%
  mutate(
    rooms = case_when(
      rooms == "More than 10" ~ 10L,                 # Convert "more than 10" to integer 10
      TRUE ~ as.integer(rooms)                      # Otherwise, convert to integer
    )
  )

df[df$rooms >= 6,]
df[!df$ads_id %in% c(95078151,98657827,98910302,99729522,99760048,100065541,100013966,100303122),]
```
Interestingly, we see quite a varying number of rooms ranging from 1 to 10, usually a unit is available in 2 - 5 rooms, hence let us investigate the other values.
From the rows, we could see that some units have 6 or more rooms despite only having size less than 2000 sq. ft., hence we will exclude ads_id %in$ c(95078151,98657827,98910302,99729522,99760048,100065541,100013966,100303122)

For the column *parking*, we decide to use kNN imputation.

<h3> Feature Engineering </h3>
Insert the ```library(ggmap)``` to get the distance from porperty address to the famous location locate in KL.

```{r}
df$full_location <- paste0(df$location, ",", df$prop_name)

# Get the unique values (Reduce the computation resource)
unique_locations <- unique(df$full_location)

# Print the unique locations
head(unique_locations) #3019 rows
```


Below is the R code, that connecting with the Google Routes API. Setup with the Google Maps library ggmap, then set up the destination location starting from property location (unique_locations). After that, using expand_grid to combine all of the location in total will be 9057 (3*3019) then running by for-loop looping all of the property location by using mapdist which return the distance between the two location in a driving model. The result will be save into a list, after that using bind_rows combine all of the inner element (data frame) into a final table. Is important to setyp the try-catch and system sleep to handle the error such as location can't be found then directly pause the for-loop function and using system sleep to reduce to overwhelmed due to lack data calling from API.

```R
R code of finding the distance from the two location.

library(ggmap)
library(dplyr)
library(dplyr)
library(tidyr) # For expand_grid
register_google(key = "API KEY") #Google Map API Key

#Set the destination location for the KL hotspot
destination_addresses <- c(
  "Kuala Lumpur, TRX",
  "Kuala Lumpur City Centre",
  "Bukit Bintang, Kuala Lumpur"
)

all_od_pairs <- expand_grid(
  origin = unique_locations,
  destination = destination_addresses
)

results_list <- list()

for (i in 1:nrow(all_od_pairs)) {
  current_origin <- all_od_pairs$origin[i]
  current_destination <- all_od_pairs$destination[i]
  
  print(paste0("Calculating route from '", current_origin, "' to '", current_destination, "'"))
  tryCatch({
     route_info_pair <- mapdist(from = current_origin, to = current_destination, mode = "driving")
     results_list[[i]] <- route_info_pair
     Sys.sleep(0.5) # Pause for 0.5 seconds to avoid rate limits (adjust as needed)
  }, 
  error = function(e) #In case, the location can't be found
  {
    message("Error calculating route: ", e$message)
    results_list[[i]] <- data.frame(from = current_origin, to = current_destination, km = NA, 
                                   hours = NA, error = e$message)
  })
 }

 final_route_df <- bind_rows(results_list)
 print(final_route_df)
```

Saving final_route_df into "property_route.csv", then read saving into data frame. Each time calling the API will increase the time consumption and computer computation resource, by saving into the csv file able to reduce it.
```{r}
df_property_route <- read.csv('property_route.csv')
head(df_property_route)
```


```{r}
df_property_distance <- df_property_route %>%
  select(from, to, km) %>% #Select the key attribute (from, to, km)
  pivot_wider(             #Splitting the km between the two location to each column by using pivot_wider
    names_from = to,       
    values_from = km,     
    names_prefix = "distance_to_" 
  )

#Rename the columns
df_property_distance <- df_property_distance %>%
  rename(
    distance_to_TRX = `distance_to_Kuala Lumpur, TRX`,
    distance_to_KLCC = `distance_to_Kuala Lumpur City Centre`,
    distance_to_BukitBintang = `distance_to_Bukit Bintang, Kuala Lumpur`
  )

# Mapping the distance into the "df" data frame, searching the matching full_location with the "from"
df <- df %>%
  left_join(df_property_distance, by = c("full_location" = "from"))

head(df)
```

To get the avearge distance, we will using rowMeans() function to calculate the distance from each row then result in a popular_location_distance column.
```{r}
df <- df %>%
  mutate(
    Popular_location_distance = rowMeans( #Calculate the mean for each distance (row)
      select(., distance_to_TRX, distance_to_KLCC, distance_to_BukitBintang),
      na.rm = TRUE #Skipping the empty value, select valid numberic value
    )
  )
head(df)
```
