---
title: "WQD7004 2024/2025 Sem2 - Group G03"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
date: "2025-06-01"
---

```{=html}
<style>
body {
  background-color: #f7cac9;
}
</style>
```

::: {style="text-align: right;"}
Group G03

Richie Teoh (24088171)\
Elmer Lee (2408236)\
Micole Chung Syn Tung (24073625)\
Lee Re Xuan ()\
Angeline Tan Jie Lin ()

01 June 2025
:::

------------------------------------------------------------------------

# **1 Objectives**

## Title:

-   Prediction of Rental Price via Regression and Classification of Property Furnishing Level in Klang Valley, Malaysia.

## Project Goal:

-   There are two (2) goals we wish to achieve with this project:

1)  *Prediction of Rental Prices*: We aim to estimate the rental price of a property based on factors like location, size, amenities. An accurate estimate will help potential renters such as students and workers to get a fair market value.

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(reshape2)
library(caret)
library(xgboost)
library(lightgbm)
```

```{r}
df <- read.csv('mudah-apartment-kl-selangor.csv')
head(df)
```

```{r}
metadata <- tibble(
  Column_Name = c(
    "ads_id",
    "prop_name",
    "completion_year",
    "monthly_rent",
    "location",
    "property_type",
    "rooms",
    "parking",
    "bathroom",
    "size",
    "furnished",
    "facilities",
    "additional_facilities"
  ),
  Description = c(
    "Unique identifier for the listing",
    "Name of the building/property",
    "Completion/established year of the property",
    "Monthly rent in Ringgit Malaysia (RM)",
    "Property location within Kuala Lumpur region",
    "Type of property (apartment, condo, flat, etc.)",
    "Number of rooms in the unit",
    "Number of parking spaces for the unit",
    "Number of bathrooms in the unit",
    "Total area of the unit in square feet",
    "Furnishing status of the unit (fully, partial, non-furnished)",
    "Main facilities available",
    "Additional facilities (e.g., proximity to mall, school, railway, etc.)"
  )
)

# Print the metadata table
metadata
```

From the metadata extracted from Kaggle above: - ads_id should be dropped as it is a unique identifier for the Mudah rental listing. - prop_name can influence the rental price, for example properties under well-known developers such as Arte tend to be higher-priced; however, it does not generalize well and unseen property names might cause issues in the model. Henceforth, we decide to drop this variable as well. - completion_year includes some NA values, will investigate further, completion_year contributes to the price and the info are usually available online, hence we will impute it manually from data found online. - monthly_rent is saved as chr type due to the "RM" prefixes, will perform appropriate data transformation. - size is saved as chr type due to the "sq. ft." suffixes, will perform appropriate data transformation.

## Data Pre-processing

1)  Filter only "serviced residences" and "condo", reduce noises for model, capturing only condo data.
2)  Remove Duplicates by ads_id, keeping only the row that contains least NA across the columns

```{r}
# Data Pre-processing

# Filter only serviced residences and condo
df <- df %>%
  filter(.data[['property_type']] %in% c('Service Residence','Condominium'))

# Check and Remove duplicated values.
duplicates_only <- df %>%
  filter(duplicated(df) | duplicated(df, fromLast = TRUE)) %>%
  arrange(ads_id)

duplicates_ads_id <- df %>%
  filter(duplicated(df['ads_id']) | duplicated(df['ads_id'], fromLast = TRUE)) %>%
  arrange(ads_id)

head(duplicates_only)
head(duplicates_ads_id)
```

```{r}
# Removing Duplicates
cleaned_df <- df %>%
  mutate(na_count = rowSums(is.na(.))) %>%
  group_by(ads_id) %>%
  arrange(na_count) %>%
  slice(1) %>%
  ungroup() %>%
  select(-na_count)

sum(duplicated(cleaned_df['ads_id']))
df <- cleaned_df

# 
```

## Check for Missing Data and Imputation

```{r}
sum(is.na(df))
colSums(is.na(df))
```

## Data Cleaning

```{r}
# Data Cleaning

# 1) Removing prefix 'RM' and suffix 'per month' from monthly rent, removing white spaces from 1 700, and convert to numeric type.
df <- df %>%
  mutate(
    monthly_rent = str_remove_all(monthly_rent, "RM|per month|,"),
    monthly_rent = str_remove_all(monthly_rent, " "),
    monthly_rent = as.numeric(monthly_rent)
  )

# 2) Removing suffix 'sq. ft.' from size, removing any white spaces, and convert to numeric type.
df <- df%>%
  mutate(
    size = str_remove_all(size, "sq\\.ft\\."),
    size = str_remove_all(size, " "),
    size = as.numeric(size)
  )

unique(df[is.na(df$completion_year), "prop_name"])
df[is.na(df$monthly_rent),]

df <- df %>%
  filter(!is.na(monthly_rent))
```

Interesting case, Affina Bay is located in Butterworth but included in this webscrapped Klang Valley Mudah listing. Henceforth, we will take more time to clean the data properly based on the listing's location.

```{r}
unique(df['location'])
```

Going through the unique values, we see a few interesting ones such as Selangor - 389, Selangor - 360, Kuala Lumpur - Others, Selangor - 517, Selangor - 639

```{r}
suspicious_location <- c('Selangor - 389', 'Selangor - 360', 'Kuala Lumpur - Others', 'Selangor - 517', 'Selangor - 639')

df[df$location %in% suspicious_location,]

df %>%
  filter(grepl("Ayuman", prop_name))
```

```{r}
# Investigating number of rooms
unique(df$rooms)

df <- df %>%
  mutate(
    rooms = case_when(
      rooms == "More than 10" ~ 10L,                 # Convert "more than 10" to integer 10
      TRUE ~ as.integer(rooms)                      # Otherwise, convert to integer
    )
  )

df[df$rooms >= 6,]
df[!df$ads_id %in% c(95078151,98657827,98910302,99729522,99760048,100065541,100013966,100303122),]
```

Interestingly, we see quite a varying number of rooms ranging from 1 to 10, usually a unit is available in 2 - 5 rooms, hence let us investigate the other values. From the rows, we could see that some units have 6 or more rooms despite only having size less than 2000 sq. ft., hence we will exclude ads_id %in\$ c(95078151,98657827,98910302,99729522,99760048,100065541,100013966,100303122)

For the column *parking*, we decide to fill the NA values with 0.

```{r}
df$parking[is.na(df$parking)] <- 0
```

For the column *completion_year*, we decide to fill NA values via median value of the locations.

```{r}
df <- df %>%
  group_by(location) %>%
  mutate(completion_year = ifelse(is.na(completion_year), median(completion_year, na.rm = TRUE), completion_year))

df_manual_prop <- read.csv('manual_prop_year.csv', header=FALSE)
colnames(df_manual_prop) <- c('prop_name','completion_year')
df <- df %>%
  left_join(df_manual_prop, by = "prop_name", suffix = c("_raw", "_manual")) %>%
  mutate(completion_year = coalesce(completion_year_raw, completion_year_manual)) %>%
  select(-completion_year_raw, -completion_year_manual)

global_median_year <- median(df$completion_year, na.rm = TRUE)

df <- df %>%
  mutate(completion_year = coalesce(completion_year, global_median_year))
```
## Final results after Cleaning
```{r}
colSums(is.na(df))
```


## Feature Engineering

Insert the `library(ggmap)` to get the distance from porperty address to the famous location locate in KL.

```{r}
df$full_location <- paste0(df$location, ",", df$prop_name)

# Get the unique values (Reduce the computation resource)
unique_locations <- unique(df$full_location)

# Print the unique locations
head(unique_locations) #3019 rows
```

Below is the R code, that connecting with the Google Routes API. Setup with the Google Maps library ggmap, then set up the destination location starting from property location (unique_locations). After that, using expand_grid to combine all of the location in total will be 9057 (3\*3019) then running by for-loop looping all of the property location by using mapdist which return the distance between the two location in a driving model. The result will be save into a list, after that using bind_rows combine all of the inner element (data frame) into a final table. Is important to setyp the try-catch and system sleep to handle the error such as location can't be found then directly pause the for-loop function and using system sleep to reduce to overwhelmed due to lack data calling from API.

``` r
R code of finding the distance from the two location.

library(ggmap)
library(dplyr)
library(tidyr) # For expand_grid
register_google(key = "API KEY") #Google Map API Key

#Set the destination location for the KL hotspot
destination_addresses <- c(
  "Kuala Lumpur, TRX",
  "Kuala Lumpur City Centre",
  "Bukit Bintang, Kuala Lumpur"
)

all_od_pairs <- expand_grid(
  origin = unique_locations,
  destination = destination_addresses
)

results_list <- list()

for (i in 1:nrow(all_od_pairs)) {
  current_origin <- all_od_pairs$origin[i]
  current_destination <- all_od_pairs$destination[i]
  
  print(paste0("Calculating route from '", current_origin, "' to '", current_destination, "'"))
  tryCatch({
     route_info_pair <- mapdist(from = current_origin, to = current_destination, mode = "driving")
     results_list[[i]] <- route_info_pair
     Sys.sleep(0.5) # Pause for 0.5 seconds to avoid rate limits (adjust as needed)
  }, 
  error = function(e) #In case, the location can't be found
  {
    message("Error calculating route: ", e$message)
    results_list[[i]] <- data.frame(from = current_origin, to = current_destination, km = NA, 
                                   hours = NA, error = e$message)
  })
 }

 final_route_df <- bind_rows(results_list)
 print(final_route_df)
```

Saving final_route_df into "property_route.csv", then read saving into data frame. Each time calling the API will increase the time consumption and computer computation resource, by saving into the csv file able to reduce it.

```{r}
df_property_route <- read.csv('property_route.csv')
head(df_property_route)
```

Searching from the missing distance between the two locations, manually adding the location.

```{r}
print(paste0("Missing value:", colSums(is.na(df_property_route))))

df_missing_route <- which(apply(is.na(df_property_route), 1, any))
df_property_route[df_missing_route,]

df_insert_missing_from <- data.frame(
  from = c("Kuala Lumpur - Kepong,Greenview Apartments", "Kuala Lumpur - Kepong,", 
           "Kuala Lumpur - Kepong,Aman Dua",
           "Kuala Lumpur - Kepong,Saujana Ria Aparment",
           "Kuala Lumpur - Kepong,Sri Saujana Apartment (Wangsa Permai)",
           "Kuala Lumpur - Kepong,VIM 3 @ Desa Park North",
           "Kuala Lumpur - Kepong,opal flat"),
  distance_to_TRX = c(12.6, 16.2, 20.2, 20.8, 20.8, 17.7, 18.2),
  distance_to_KLCC = c(9.9, 13.5, 16.9, 17.6, 17.6, 15.2, 14.7),
  distance_to_BukitBintang = c(11.8, 15.5, 19.4, 20.1, 20.1, 16.9, 16.6)
)
df_insert_missing_from
```

Storing the searching route result into main data frame (df).
```{r}
df_property_distance <- df_property_route %>%
  select(from, to, km) %>% #Select the key attribute (from, to, km)
  pivot_wider(             #Splitting the km between the two location to each column by using pivot_wider
    names_from = to,       
    values_from = km,     
    names_prefix = "distance_to_" 
  )

#Rename the columns
df_property_distance <- df_property_distance %>%
  rename(
    distance_to_TRX = `distance_to_Kuala Lumpur, TRX`,
    distance_to_KLCC = `distance_to_Kuala Lumpur City Centre`,
    distance_to_BukitBintang = `distance_to_Bukit Bintang, Kuala Lumpur`
  )

#Checking the missing value before implement replacement
print(paste0("Before Implement Replacement for Missing value:", colSums(is.na(df_property_distance))))

df_property_distance_new <- data.frame(df_property_distance) #Make a copy

#Drop all of the NA value
df_property_distance_new <- df_property_distance_new %>%
  drop_na(distance_to_TRX, distance_to_KLCC, distance_to_BukitBintang)

#Merge the replacement data frame with the copy of property distance data frame
df_property_distance <- bind_rows(df_property_distance_new, df_insert_missing_from)

#Check missing value after replacement
print(paste0("After Implement Replacement for Missing value:", colSums(is.na(df_property_distance))))

# Mapping the distance into the "df" data frame, searching the matching full_location with the "from"
df <- df %>%
  left_join(df_property_distance, by = c("full_location" = "from"))

head(df)
```

To get the average distance, we will using rowMeans() function to calculate the distance from each row then result in a popular_location_distance column.

```{r}
df_import_PopularLocation <- data.frame(df) #Make a copy

df_import_PopularLocation <- df_import_PopularLocation %>%
  mutate(
    Popular_location_distance = rowMeans( #Calculate the mean for each distance (row)
      select(., distance_to_TRX, distance_to_KLCC, distance_to_BukitBintang),
      na.rm = TRUE #Skipping the empty value, select valid numberic value
    )
  )
df <- df_import_PopularLocation
head(df)
```

```{r}
df <- df %>%
  mutate(near_public_transport = as.integer(grepl("Near KTM/LRT", additional_facilities, ignore.case = TRUE)))
head(df)
```


# EDA
```{r}
df_EDA = data.frame(df) #Make a copy for EDA
head(df_EDA)
```

## Univariate Analysis

### Histogram of Monthly Rent
This section shows a histogram of monthly rent. It is split into > 5,000 (MYR) and < 5,000 (MYR) parts. Based on the summary results, the upper bound is about 2,100 (MYR), therefore splitting into parts below 5,000 (MYR) able provides a clearer visualization.
```{r}
hist_data <- (df_EDA$monthly_rent)
summary(hist_data) #Show the summary

hist_data_below <- hist_data[hist_data < 5000]

hist(
  hist_data_below, xlab = "Monthly Accommodation Rent (MYR)", main = "Monthly Accommodation Rent Distribution (< 5k)",
  col = "#A8E4A0", border = "black", breaks = 40
)

```
### Histogram of Property Size
This section shows a histogram of size. It is split into > 3,000 (sq ft) and < 3,000 (sq ft) parts. Based on the summary results, the upper bound is about 1,100 (sq ft), therefore splitting into parts below 3,000 (sq ft) can provides a clearer visualization.
```{r}
hist_data_size <- (df_EDA$size)
summary(hist_data_size) #Show the summary

hist_data_size_below <- hist_data_size[hist_data_size < 3000]

hist(
  hist_data_size_below, xlab = "Property Size (sq ft)", main = "Property Size Distribution (< 3,000)",
  col = "#A8E4A0", border = "black", breaks = 30
)

```
### Bar Chart
This section plot horizontal bar graphs to illustrate the frequency/count of each nominal or ordinal data.
```{r}
plot_bar_chart_char <- function(bar_chart_columns) {
  
  for (col in bar_chart_columns) {
    counts_data <- df_EDA[[col]]
  
    unique(df_EDA$col)
    
    # Count occurrences
    category_counts <- table(counts_data)
    
    # Plot horizontal bar chart
    barplot(category_counts,
            col = "skyblue",
            main = paste("Accommodation" , col, "Distribution"),
            xlab = "Count")
  }
}

bar_chart_columns <- c("parking", "rooms", "bathroom", "completion_year")
plot_bar_chart_char(bar_chart_columns)

```
### Pie Chart of Property Type
```{r}
property_data <- df_EDA %>% count(property_type)

ggplot(property_data, aes(x = "", y = n, fill = property_type)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(n / sum(n) * 100, 1), "%")), position = position_stack(vjust = 0.5)) + 
  labs(title = "Property Type") +
  theme(legend.title = element_blank())
```

### Pie Chart of Near Public Transport
```{r}
near_public_transport_data <- df_EDA %>% count(near_public_transport)

near_public_transport_data$near_public_transport[near_public_transport_data$near_public_transport == 0] <- "No"
near_public_transport_data$near_public_transport[near_public_transport_data$near_public_transport == 1] <- "Yes"

ggplot(near_public_transport_data, aes(x = "", y = n, fill = near_public_transport)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(n / sum(n) * 100, 1), "%")), position = position_stack(vjust = 0.5)) + 
  labs(title = "Near Public Transport") +
  theme(legend.title = element_blank())
```

### Pie Chart of Near Public Transport
```{r}
furnished_data <- df_EDA %>% count(furnished)

ggplot(furnished_data, aes(x = "", y = n, fill = furnished)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(n / sum(n) * 100, 1), "%")), position = position_stack(vjust = 0.5)) + 
  labs(title = "Property Type") +
  theme(legend.title = element_blank())
```

## Multivariate Analysis

### BoxPlot - Two Relationship
In this section, we will use a scatter plot to understand the relationship between monthly rent and property size. The scatter plot is filtered based on the upper range to provide a clear visualization.
```{r}
#Define the bin size 
breaks <- seq(0, 3000, by = 500)
rent_size_label <- paste(head(breaks, -1), tail(breaks, -1), sep = "–") #Set the rent_size range ([0],[1])

# Filter the size < 3000 and rent < 5000
boxplot_data_filter <- df_EDA %>%
  filter(size < 3000, monthly_rent < 5000) %>%
  mutate(property_size_bin = cut(size, breaks = breaks, labels = rent_size_label))

# Box plot: Size by Rent Range
ggplot(boxplot_data_filter, aes(x = property_size_bin, y = monthly_rent)) +
  geom_boxplot(fill = "#87CEFA", color = "#4682B4", outlier.alpha = 0.3) +
  labs(
    title = "Property Size Distribution by Monthly Accommodation Rent Range",
    x = "Property Size (sq ft)",
    y = "Monthly Rent Range (MYR)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 0.5))
```

In this section, we will use a scatter plot to understand the relationship between monthly rent and average popular location distance to KLCC, TRX and Bukit Bintang. The scatter plot is filtered based on the upper range to provide a clear visualization.
```{r}
# Filter the data
most_frequent_locations <- df_EDA %>%
  count(location, sort = TRUE) #Find the most frequency location

#Filter top 20 location and < 5000 money rent
boxplot_data_filter <- df_EDA %>%
  filter(monthly_rent < 5000, location %in% most_frequent_locations$location[0:20])

ggplot(boxplot_data_filter, aes(x = reorder(location, monthly_rent, median), y = monthly_rent)) +
  geom_boxplot(fill = "#87CEFA", color = "#4682B4", outlier.alpha = 0.3) +
  labs(
    title = "Monthly Accommodation Rent Distribution by Top 20 Locations",
    x = "Location",
    y = "Monthly Rent (MYR)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#Arrange completion year and < 5000 money rent
boxplot_data_filter <- df_EDA %>%
  mutate(completion_year = as.integer(completion_year)) %>%
  filter(monthly_rent < 5000) %>%
  arrange(completion_year) 

#Using factor() arrange into categories type (Grouping)
boxplot_data_filter$completion_year <- factor(
  boxplot_data_filter$completion_year
) 

ggplot(boxplot_data_filter, aes(x = completion_year, y = monthly_rent)) +
  geom_boxplot(fill = "#87CEFA", color = "#4682B4", outlier.alpha = 0.3) +
  labs(
    title = "Monthly Accommodation Rent Distribution by Year",
    x = "Year",
    y = "Monthly Rent (MYR)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Correlation Matrix
This section showcase a correlation matrix between each feature.
```{r}
heatmap_data <- df_EDA %>%
  select(where(is.numeric)) #Select only numeric data (remove such as prop_name, e.g.)

heatmap_data <- heatmap_data %>% select(-ads_id)  #Remove ads_id

corr_data <- cor(heatmap_data, use = "complete.obs") #Transfer into numeric (-1 , 1)
melted_corr <- melt(corr_data) #into a format for ggplot
heap_map_color <- colorRampPalette(c("skyblue", "white", "#cc474b"))(200)

# Plot
ggplot(melted_corr, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(value,2)), size = 3) +
  scale_fill_gradientn(colors = heap_map_color, limits = c(-1, 1)) +
  labs(title = "Correlation Matrix", x = NULL, y = NULL) + #Add NULL else will have Var2 and Var1
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.2, size = 16, face = "bold")
  )
```

# Data Modelling

As this project is focusing in 2 different targets, predicting rental prices and furnishing status, different models are used for each prediction. 

```{r}
df_model = data.frame(df) #Make a copy for Data Modelling
head(df_model)
```

##Data Standardization & Label encoding

```{r}
df_model <- df_model %>%
  mutate(
    
#Data Standardization for 'monthly_rent', 'size', and 'Popular_location_distance' columns
    
    monthly_rent_standardized = scale (monthly_rent),
    size_standardized = scale (size),
    Popular_location_distance_standardized = scale (Popular_location_distance),
    
#Label encoding for 'property_type' and 'furnished' columns

    property_type_encoded = as.integer(factor(property_type)),
    furnished_encoded = as.integer(factor(furnished))
  )

#Transform the 'completion_year' column to age of the property

current_year <- 2025
df_model$property_age <- current_year - df_model$completion_year

#One-hot encoding for 'facilities' and 'additional facilities' columns

#Combine both 'facilities' and 'additional facilities' columns
df_combined_facilities <- df_model %>%
  mutate(facilities_combined = paste(facilities, additional_facilities, sep = ", "))

#Remove unwanted characters such as 10, 11, 6 in 'facilities' column
df_combined_facilities$facilities_combined <- gsub("[^a-zA-Z, ]", "", df_combined_facilities$facilities_combined)

#Split each facility into separate rows based on the comma between
df_facilities <- df_combined_facilities %>%
  separate_rows(facilities_combined, sep = ", ")  # Split facilities by comma and space

#Replace blank spaces with underscores
df_facilities <- df_facilities %>%
  mutate(facilities_combined = gsub(" ", "_", facilities_combined)) %>%  
  filter(facilities_combined != "")  #Remove empty strings 

#Create binary code for each facility
df_facilities_encoded <- df_facilities %>%
  distinct(facilities_combined) %>%  # Make sure each facility appears only once
  pivot_wider(names_from = facilities_combined, values_from = facilities_combined, 
              values_fn = length, values_fill = list(facilities_combined = 0))  

#Merge back to the original dataframe
df_model_encoded <- bind_cols(df_model, df_facilities_encoded)


#Filter out unwanted columns
drop_columns <- c("ads_id","prop_name", "full_location", "location","property_type", "furnished", "facilities", "additional_facilities", "region",  "completion_year","distance_to_TRX", "distance_to_KLCC", "distance_to_BukitBintang", "Popular_location_distance", "size", "monthly_rent")

df_model <- df_model_encoded%>% select(-all_of(drop_columns))

head(df_model)

```

## Data Splitting

Data splitting is performed separately for both targets variables: monthly_rent for the regression model (predicting rental price) and furnished for the classification model (predicting furnishing status). 

The dataset is split into training and testing sets. Training sets are used to build model, while testing sets are used to evaluate the model performance. In this case, we split the data into 80% train set and 20% test set. Separate splits are created for each model according to their respective target variable.

### Regression Model (Predict Rental Price)

```{r}
#For reproducibility
set.seed(123) 

#Split in 80/20 for train and test sets
split = 0.80 

```

Create train and test sets using createDataPartition(). This function helps to split the data in a way that reflects the targeted variables (Monthly_rent) distribution. 

```{r}
train_index_regression <- createDataPartition(df_model$monthly_rent_standardized, p=split, list=FALSE) 

#Create training and testing datasets
X_train_regression <- df_model[train_index_regression,]%>% select( -monthly_rent_standardized)
y_train_regression <- df_model$monthly_rent_standardized[train_index_regression]

X_test_regression <- df_model[-train_index_regression,]%>% select(-monthly_rent_standardized)
y_test_regression <- df_model$monthly_rent_standardized[-train_index_regression]

# Print the shape of X_train_regression and y_train_regression
cat("X_train_regression shape:", dim(X_train_regression), "y_train_regression shape:", length(y_train_regression), "\n")

# Print the shape of X_test_regression and y_test_regression
cat("X_test_regression shape:", dim(X_test_regression), "y_test_regression shape:", length(y_test_regression))
```

### Classification Model (Predict Furnishing Status)

```{r}
#For reproducibility
set.seed(123) 

#Split in 80/20 for train and test sets
split = 0.80 

#Create train and test sets using createDataPartition
train_index_classification <- createDataPartition(df_model$furnished_encoded, p=split, list=FALSE) 

X_train_classification <- df_model[train_index_classification,]%>% select( -furnished_encoded)
y_train_classification <- df_model$furnished_encoded[train_index_classification]

X_test_classification <- df_model[-train_index_classification,]%>% select( -furnished_encoded)
y_test_classification <- df_model$furnished_encoded[-train_index_classification]

# Print the shape of X_train_classification and y_train_classification for 
cat("X_train_classification shape:", dim(X_train_classification), "y_train_classification shape:", length(y_train_classification), "\n")

# Print the shape of X_test_classification and y_test_classification
cat("X_test_classification shape:", dim(X_test_classification), "y_test_classification shape:", length(y_test_classification))
```

## Model Training

In this section, the respective training sets are used to build regression and classification models with different algorithms. 

```{r}
#Cross-validation 
train_ctrl <- trainControl(
  method = "cv",
  number = 5, 
  verboseIter   = TRUE, #Provide detailed, real-time feedback on training process
  savePredictions = "final" #To save out-of-sample predictions from the final model
)

#For Regression metrics
regression_metrics <- c("MAE", "MSE", "RMSE", "R²")

#For Classification metrics
classification_metrics <- c("Accuracy", "Precision", "Recall", "F1-score", "AUC-ROC", "Kappa")

```

### Regression Model (Predict Rental Price)

For Regression models, algorithms like Linear Regression, XGBoost, LightGBM,  Decision Tree, Random Forest, Neural Networks, and CatBoost are trained to predict the rental price based on key property features such as 

# size, location, number of rooms, and furnishing status.(nned to change)+metrics

#### Linear Regression Model

```{r}

#Train Linear Regression model
lm_model <- train (
  x = X_train_regression, 
  y = y_train_regression, 
  method = "lm", 
  trControl = train_ctrl,
  metric = "RMSE" #train() can only accept one metric at a time
)

#Prediction
lm_prediction <- predict(lm_model, X_test_regression)

#Evaluate model performance
lm_rmse <- RMSE(lm_prediction, y_test_regression)
lm_r2 <- R2(lm_prediction, y_test_regression)
lm_mae <- MAE(lm_prediction, y_test_regression)
lm_mse <- (lm_rmse)^2


cat("Linear Regression RMSE: ", lm_rmse, "\n")
cat("Linear Regression R²: ", lm_r2, "\n")
cat("Linear Regression MAE: ", lm_mae, "\n")
cat("Linear Regression MSE: ", lm_mse, "\n")

```
####XGBoost Model

```{r}

#Train XGBoost model
xgb_model <- train (
  x = X_train_regression, 
  y = y_train_regression, 
  method = "xgbTree", 
  trControl = train_ctrl,
  metric = "RMSE" #train() can only accept one metric at a time
)

#Prediction
xgb_prediction <- predict(xgb_model, X_test_regression)

#Evaluate model performance
xgb_rmse <- RMSE(xgb_prediction, y_test_regression)
xgb_r2 <- R2(xgb_prediction, y_test_regression)
xgb_mae <- MAE(xgb_prediction, y_test_regression)
xgb_mse <- (xgb_rmse)^2

cat("XGBoost RMSE: ", xgb_rmse, "\n")
cat("XGBoost R²: ", xgb_r2, "\n")
cat("XGBoost MAE: ", xgb_mae, "\n")
cat("XGBoost MSE: ", xgb_mse, "\n")

```

####LightGBM Model

```{r}
library(lightgbm)

# (LightGBM is not supported in caret)
#Convert to matrix format (as LightGBM supports)
lgb_train_data <- lgb.Dataset(data = as.matrix(X_train_regression), label = y_train_regression)

#Training contrl for LightGBM model
lgb_train_ctrl <- list(
  objective = "regression",  
  metric = "rmse",            #RMSE metric
  num_leaves = 31,            #Number of leaves in each tree
  learning_rate = 0.05,       #Learning rate
  nrounds = 100               #Number of boosting rounds

)

#Train LightGBM model 
lgb_model <- lgb.train(
  params = lgb_train_ctrl,
  data = lgb_train_data,
  nrounds = lgb_train_ctrl$nrounds
)
  
#Prediction
lgb_prediction <- predict(lgb_model, as.matrix (X_test_regression))

#Evaluate model performance
lgb_rmse <- RMSE(lgb_prediction, y_test_regression)
lgb_r2 <- R2(lgb_prediction, y_test_regression)
lgb_mae <- MAE(lgb_prediction, y_test_regression)
lgb_mse <- (lgb_rmse)^2

cat("LightGBM RMSE: ", lgb_rmse, "\n")
cat("LightGBM R²: ", lgb_r2, "\n")
cat("LightGBM MAE: ", lgb_mae, "\n")
cat("LightGBM MSE: ", lgb_mse, "\n")

```

####Decision Tree Model
```{r}

#Train Decision Tree model
dt_model <- train (
  x = X_train_regression, 
  y = y_train_regression, 
  method = "rpart", 
  trControl = train_ctrl,
  metric = "RMSE" #train() can only accept one metric at a time
)

#Prediction
dt_prediction <- predict(dt_model, X_test_regression)

#Evaluate model performance
dt_rmse <- RMSE(dt_prediction, y_test_regression)
dt_r2 <- R2(dt_prediction, y_test_regression)
dt_mae <- MAE(dt_prediction, y_test_regression)
dt_mse <- (dt_rmse)^2

cat("Decision Tree RMSE: ", dt_rmse, "\n")
cat("Decision Tree R²: ", dt_r2, "\n")
cat("Decision Tree MAE: ", dt_mae, "\n")
cat("Decision Tree MSE: ", dt_mse, "\n")

```

####Random Forest Model
```{r}

#Train Random Forest model
rf_model <- train (
  x = X_train_regression, 
  y = y_train_regression, 
  method = "rf", 
  trControl = train_ctrl,
  metric = "RMSE" #train() can only accept one metric at a time
)

#Prediction
rf_prediction <- predict(rf_model, X_test_regression)

#Evaluate model performance
rf_rmse <- RMSE(rf_prediction, y_test_regression)
rf_r2 <- R2(rf_prediction, y_test_regression)
rf_mae <- MAE(rf_prediction, y_test_regression)
rf_mse <- (rf_rmse)^2

cat("Random Forest RMSE: ", rd_rmse, "\n")
cat("Random Forest R²: ", rf_r2, "\n")
cat("Random Forest MAE: ", rf_mae, "\n")
cat("Random Forest MSE: ", rf_mse, "\n")

```

####Neural Network Model
```{r}

#Train Neural Network model
nn_model <- train (
  x = X_train_regression, 
  y = y_train_regression, 
  method = "nnet", 
  trControl = train_ctrl,
  metric = "RMSE", #train() can only accept one metric at a time
  trace = FALSE #To suppress the training output
)

#Prediction
nn_prediction <- predict(nn_model, X_test_regression)

#Evaluate model performance
nn_rmse <- RMSE(nn_prediction, y_test_regression)
nn_r2 <- R2(nn_prediction, y_test_regression)
nn_mae <- MAE(nn_prediction, y_test_regression)
nn_mse <- (nn_rmse)^2

cat("Neural Network RMSE: ", nn_rmse, "\n")
cat("Neural Network R²: ", nn_r2, "\n")
cat("Neural Network MAE: ", nn_mae, "\n")
cat("Neural Network MSE: ", nn_mse, "\n")

```

####CatBoost Model
```{r}

#Train CatBoost model
catb_model <- train (
  x = X_train_regression, 
  y = y_train_regression, 
  method = "catboost", 
  trControl = train_ctrl,
  metric = "RMSE" #train() can only accept one metric at a time
)

#Prediction
catb_prediction <- predict(catb_model, X_test_regression)

#Evaluate model performance
catb_rmse <- RMSE(catb_prediction, y_test_regression)
catb_r2 <- R2(catb_prediction, y_test_regression)
catb_mae <- MAE(catb_prediction, y_test_regression)
catb_mse <- (catb_rmse)^2

cat("CatBoost RMSE: ", catb_rmse, "\n")
cat("CatBoost R²: ", catb_r2, "\n")
cat("CatBoost MAE: ", catb_mae, "\n")
cat("CatBoost MSE: ", catb_mse, "\n")

```

### Classification Model (Predict Furnishing Status)

For classification models, algorithms like XGBoost, LightGBM, Logistic Regression, CatBoost, Decision Trees, Random Forest Classifier, Support Vector Machines (SVM), and Neural Network are trained to predict the furnishing status of the property based on property features such as 

#size, location, number of bedrooms, and rental price.(need to change) +metrics

####XGBoost Model
```{r}

#Train XGBoost model
xgb_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "xgbTree", 
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
xgb_class_prediction <- predict(xgb_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
xgb_class_cm <- confusionMatrix(xgb_class_prediction, y_test_classification)
print (xgb_class_cm)

```

####LightGBM Model
```{r}

#Train XGBoost model
lgb_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "lightgbm", 
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
lgb_class_prediction <- predict(lgb_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
lgb_class_cm <- confusionMatrix(lgb_class_prediction, y_test_classification)
print (lgb_class_cm)

```

####Logistic Regression Model
```{r}

#Train XGBoost model
log_reg_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "glm", 
  family = "binomial", #To tell the model that the target variable is binary
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
log_reg_class_prediction <- predict(log_reg_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
log_reg_class_cm <- confusionMatrix(log_reg_class_prediction, y_test_classification)
print (log_reg_class_cm)

```

####CatBoost Model
```{r}

#Train CatBoost model
catb_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "catboost", 
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
catb_class_prediction <- predict(catb_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
catb_class_cm <- confusionMatrix(catb_class_prediction, y_test_classification)
print (catb_class_cm)

```

####Decision Tree Classifier
```{r}

#Train Decision Tree Classifier
dt_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "rpart", 
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
dt_class_prediction <- predict(dt_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
dt_class_cm <- confusionMatrix(dt_class_prediction, y_test_classification)
print (dt_class_cm)

```

####Random Forest Classifier
```{r}

#Train Random Forest Classifier
rf_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "rf", 
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
rf_class_prediction <- predict(rf_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
rf_class_cm <- confusionMatrix(rf_class_prediction, y_test_classification)
print (rf_class_cm)

```

####Support Vector Machines (SVM) Classifier
```{r}

#Train Support Vector Machines (SVM) Classifier
svm_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "svmRadial", 
  trControl = train_ctrl,
  metric = "Accuracy" #train() can only accept one metric at a time
)

#Prediction
svm_class_prediction <- predict(svm_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
svm_class_cm <- confusionMatrix(svm_class_prediction, y_test_classification)
print (svm_class_cm)

```

####Neural Network Classifier
```{r}

#Train Neural Network Classifier 
nn_class_model <- train (
  x = X_train_classification, 
  y = y_train_classification, 
  method = "nnet", 
  trControl = train_ctrl,
  metric = "Accuracy", #train() can only accept one metric at a time
  trace = FALSE #To suppress the training output
)

#Prediction
nn_class_prediction <- predict(nn_class_model, X_test_classification)

#Evaluate model performance - Confusion Matrix
nn_class_cm <- confusionMatrix(nn_class_prediction, y_test_classification)
print (nn_class_cm)

```